---
title: "Tissue pO~2~ and Neuronal Death in rTg4510 vs Wild-Type Mice"
subtitle: "Mixed-effects logistic regression with rare-event and clustering safeguards"
author: "Nancy E. Ruiz-Uribe"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    theme: flatly
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 5, dpi = 150, fig.align = "center"
)
```

# 1. Configuration

Edit the two paths below; everything downstream is path-independent.

```{r config}
TG_FILE <- "data/Tg_results.xlsx"   # rTg4510
WT_FILE <- "data/WT_results.xlsx"   # wild-type

# Colour scheme (consistent across all figures)
COL_TG    <- "#E87461"   # transgenic  (rTg4510)
COL_WT    <- "#6497B1"   # wild-type
COL_DYING <- "#B22222"   # dying neurons
COL_ALIVE <- "#9AA5AD"   # non-dying neurons

PO2_UNIT  <- 10          # report odds ratios per 10 mmHg
SEED      <- 20260904
set.seed(SEED)
```

```{r packages}
suppressPackageStartupMessages({
  library(readxl)      # read the two workbooks
  library(dplyr)       # data manipulation
  library(tidyr)
  library(ggplot2)     # figures
  library(lme4)        # glmer  -- primary model
  library(broom.mixed) # tidy() for merMod
  library(brglm2)      # Firth penalised logistic -- rare-event safeguard
  library(geepack)     # GEE with mouse as cluster -- robustness check
  library(knitr)
})

theme_set(
  theme_classic(base_size = 12) +
    theme(
      strip.background = element_blank(),
      strip.text       = element_text(face = "bold"),
      legend.position  = "top",
      plot.title       = element_text(face = "bold")
    )
)
```

# 2. Data import and construction of the within-mouse predictor

The two workbooks share the same schema: `CaseID` (mouse), `NeuronID`,
`pO2` (mmHg), `z`, `isDying` (logical), `pO2c`.

`NeuronID` is only unique **within** a mouse, so the neuron key is
`CaseID` + `NeuronID`.

```{r import}
read_one <- function(path, genotype) {
  read_excel(path, sheet = 1) |>
    mutate(
      Genotype = genotype,
      CaseID   = as.character(CaseID),
      isDying  = as.logical(isDying)
    )
}

dat_raw <- bind_rows(
  read_one(TG_FILE, "Tg"),
  read_one(WT_FILE, "WT")
)

glimpse(dat_raw)
```

## 2.1 Recompute the centred predictors

Baseline pO~2~ differs several-fold between mice, so the model uses
**within-mouse centred** pO~2~: each neuron's deviation from its own
mouse's mean. This asks the biologically pointed question — *within a
given mouse, do dying neurons sit in lower-pO~2~ tissue than that same
mouse's surviving neurons?* — and is immune to between-animal baseline
differences, imaging session, and depth.

We recompute rather than trust the supplied columns, then verify they
agree.

```{r center}
dat <- dat_raw |>
  group_by(CaseID) |>
  mutate(
    pO2_case_mean = mean(pO2),
    pO2_case_sd   = sd(pO2),
    pO2c_calc     = pO2 - pO2_case_mean,               # mmHg from mouse mean
    z_calc        = (pO2 - pO2_case_mean) / pO2_case_sd # within-mouse SD units
  ) |>
  ungroup() |>
  mutate(
    Genotype  = factor(Genotype, levels = c("WT", "Tg")),  # WT = reference
    Dying     = factor(isDying, levels = c(FALSE, TRUE),
                       labels = c("Non-dying", "Dying")),
    pO2c_10   = pO2c_calc / PO2_UNIT,   # per-10-mmHg within-mouse deviation
    pO2_10    = pO2 / PO2_UNIT          # per-10-mmHg absolute (sensitivity)
  )

# Verify our centring reproduces the supplied columns
tibble(
  check          = c("pO2c", "z"),
  max_abs_diff   = c(max(abs(dat$pO2c - dat$pO2c_calc)),
                     max(abs(dat$z    - dat$z_calc)))
) |> kable(digits = 10, caption = "Agreement with supplied columns")
```

## 2.2 Data audit

```{r audit}
stopifnot(
  !any(is.na(dat$pO2)),
  !any(is.na(dat$isDying)),
  !anyDuplicated(dat[, c("CaseID", "NeuronID")])
)

per_mouse <- dat |>
  group_by(Genotype, CaseID) |>
  summarise(
    n_neurons   = n(),
    n_dying     = sum(isDying),
    pct_dying   = 100 * mean(isDying),
    pO2_mean    = mean(pO2),
    pO2_sd      = sd(pO2),
    pO2_dying   = ifelse(n_dying > 0, mean(pO2[isDying]), NA_real_),
    pO2_alive   = mean(pO2[!isDying]),
    delta_pO2c  = ifelse(n_dying > 0,
                         mean(pO2c_calc[isDying]) - mean(pO2c_calc[!isDying]),
                         NA_real_),
    .groups = "drop"
  )

kable(per_mouse, digits = 2,
      caption = "Per-mouse summary. delta_pO2c = mean within-mouse centred pO2 of dying minus non-dying neurons (negative = dying neurons in relatively hypoxic tissue).")
```

```{r event-count}
cat("Total neurons:", nrow(dat),
    "| Dying:", sum(dat$isDying),
    sprintf("(%.2f%%)\n", 100 * mean(dat$isDying)))
with(dat, table(Genotype, Dying)) |> kable(caption = "Events by genotype")
```

> **Events-per-variable warning.** There are only `r sum(dat$isDying)`
> dying neurons in total. The usual rule of thumb is ~10 events per
> estimated parameter, so the interaction model (3 fixed slopes +
> variance component) is already at the edge of what these data
> support. This is why we fit three estimators and a permutation test
> rather than reporting a single p-value.

# 3. Exploratory figures

## 3.1 Raw pO~2~ per mouse, with dying neurons overlaid

```{r fig-raw, fig.height=5.5}
ggplot(dat, aes(x = reorder(CaseID, pO2_case_mean), y = pO2)) +
  geom_violin(aes(fill = Genotype), colour = NA, alpha = 0.30,
              scale = "width", trim = TRUE) +
  geom_jitter(data = filter(dat, !isDying),
              width = 0.14, height = 0, size = 0.7,
              colour = COL_ALIVE, alpha = 0.35) +
  geom_jitter(data = filter(dat, isDying),
              width = 0.10, height = 0, size = 2.4,
              colour = COL_DYING, shape = 18) +
  stat_summary(fun = mean, geom = "point", shape = 95, size = 8,
               colour = "black") +
  scale_fill_manual(values = c(WT = COL_WT, Tg = COL_TG)) +
  facet_grid(~ Genotype, scales = "free_x", space = "free_x") +
  labs(
    title = "Absolute tissue pO2 varies several-fold between mice",
    subtitle = "Red diamonds = dying neurons; grey = non-dying; black bar = mouse mean",
    x = NULL, y = expression(pO[2]~"(mmHg)")
  ) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))
```

This figure is the argument for within-mouse centring: mouse means span
roughly `r round(min(per_mouse$pO2_mean))`–`r round(max(per_mouse$pO2_mean))` mmHg,
which would otherwise dominate any absolute-pO~2~ comparison.

## 3.2 Within-mouse centred pO~2~ by outcome

```{r fig-centered}
ggplot(dat, aes(x = Dying, y = pO2c_calc, fill = Genotype)) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey55") +
  geom_violin(colour = NA, alpha = 0.35, scale = "width") +
  geom_jitter(width = 0.12, height = 0, size = 0.8, alpha = 0.35,
              colour = COL_ALIVE) +
  stat_summary(fun.data = mean_cl_boot, geom = "pointrange",
               size = 0.6, colour = "black") +
  scale_fill_manual(values = c(WT = COL_WT, Tg = COL_TG)) +
  facet_wrap(~ Genotype) +
  labs(
    title = "Within-mouse centred pO2 in dying vs non-dying neurons",
    subtitle = "Each neuron expressed as deviation from its own mouse's mean; black = mean +/- bootstrap 95% CI",
    x = NULL, y = expression(Delta*pO[2]~"from mouse mean (mmHg)")
  )
```

# 4. Primary model — mixed-effects logistic regression

$$
\text{logit}\;\Pr(\text{isDying}_{ij}=1)
= \beta_0 + \beta_1\,\Delta pO_{2,ij}
+ \beta_2\,\text{Tg}_j
+ \beta_3\,(\Delta pO_{2,ij}\times \text{Tg}_j)
+ u_j,\qquad u_j\sim N(0,\sigma_u^2)
$$

for neuron $i$ in mouse $j$. The random intercept $u_j$ lets each mouse
have its own baseline death rate (they range from
`r sprintf('%.2f%%', min(per_mouse$pct_dying))` to
`r sprintf('%.2f%%', max(per_mouse$pct_dying))`), which is what makes the
neuron-level observations exchangeable within a mouse.

```{r glmm}
m_glmm <- glmer(
  isDying ~ pO2c_10 * Genotype + (1 | CaseID),
  data    = dat,
  family  = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa",
                         optCtrl = list(maxfun = 2e5))
)

summary(m_glmm)
```

## 4.1 Convergence and singularity diagnostics

With only 7 clusters the variance component is hard to estimate; check
before interpreting.

```{r glmm-diag}
cat("Singular fit:", isSingular(m_glmm, tol = 1e-4), "\n")
cat("Random-intercept SD (logit scale):",
    round(as.data.frame(VarCorr(m_glmm))$sdcor, 4), "\n\n")

# Refit with several optimisers; fixed effects should be stable
allfits <- allFit(m_glmm, verbose = FALSE)
summary(allfits)$fixef |> round(3) |>
  kable(caption = "Fixed effects across optimisers (should agree to ~2 decimals)")
```

> **On a singular fit.** With these data `lme4` estimates the
> between-mouse variance at essentially **zero** once `Genotype` is in
> the model — i.e. genotype already explains the differences in
> baseline death rate across the 7 animals, leaving no residual
> mouse-level variance to estimate. This is a boundary estimate, not a
> convergence failure, and the fixed effects remain valid. It does mean
> the random intercept is doing no work, so the **Firth model in §5.1**
> — which removes each mouse's baseline exactly, as a fixed effect
> rather than a draw from a normal distribution — is the more
> defensible specification for the record. Report both.

## 4.2 Odds ratios

```{r glmm-or}
or_glmm <- broom.mixed::tidy(m_glmm, effects = "fixed",
                             conf.int = TRUE, exponentiate = TRUE) |>
  select(term, OR = estimate, conf.low, conf.high, p.value)

kable(or_glmm, digits = 3,
      caption = paste0("GLMM odds ratios (pO2 terms are per ", PO2_UNIT,
                       " mmHg of within-mouse deviation)"))
```

**How to read this table**

| Term | Question it answers |
|---|---|
| `pO2c_10` | In **WT**, how do the odds of a neuron being dying change per 10 mmHg *above* its mouse's mean pO~2~? OR < 1 means dying neurons sit in relatively hypoxic tissue. |
| `GenotypeTg` | Baseline odds of death in Tg vs WT at each mouse's own mean pO~2~ (with only 7 mice this is imprecise — it is essentially a 4-vs-3 comparison). |
| `pO2c_10:GenotypeTg` | **The interaction.** Does the pO~2~–death relationship *differ* between Tg and WT? This is the headline test. |

## 4.3 Genotype-specific slopes

The `pO2c_10` coefficient is the WT slope only; the Tg slope is the sum
of the main effect and the interaction. Extract both with their
uncertainty.

```{r slopes}
b  <- fixef(m_glmm)

# Build contrasts by NAME, not position, so this stays correct if the
# model formula changes.
mk <- function(terms) {
  v <- setNames(numeric(length(b)), names(b))
  v[terms] <- 1
  v
}
L <- rbind(
  WT = mk("pO2c_10"),
  Tg = mk(c("pO2c_10", "pO2c_10:GenotypeTg"))
)
V  <- as.matrix(vcov(m_glmm))

est <- as.vector(L %*% b)
se  <- sqrt(diag(L %*% V %*% t(L)))

slopes_tbl <- tibble(
  Genotype = rownames(L),
  logOR    = est,
  SE       = se,
  OR       = exp(est),
  conf.low = exp(est - 1.96 * se),
  conf.high= exp(est + 1.96 * se),
  z        = est / se,
  p.value  = 2 * pnorm(-abs(est / se))
)

kable(slopes_tbl, digits = 3,
      caption = paste0("Genotype-specific pO2 slopes, per ", PO2_UNIT,
                       " mmHg within-mouse deviation"))
```

## 4.4 Likelihood-ratio tests

```{r lrt}
m_no_int  <- update(m_glmm, . ~ pO2c_10 + Genotype + (1 | CaseID))
m_no_pO2  <- update(m_glmm, . ~ Genotype + (1 | CaseID))

anova(m_no_int, m_glmm) |>
  kable(digits = 4, caption = "LRT: does the pO2 effect differ by genotype?")

anova(m_no_pO2, m_no_int) |>
  kable(digits = 4, caption = "LRT: is there any pO2 effect (common slope)?")
```

## 4.5 Predicted probability curve

```{r fig-pred, fig.height=4.8}
newd <- expand_grid(
  pO2c_10  = seq(min(dat$pO2c_10), max(dat$pO2c_10), length.out = 200),
  Genotype = factor(c("WT", "Tg"), levels = c("WT", "Tg"))
)

X    <- model.matrix(~ pO2c_10 * Genotype, newd)
eta  <- as.vector(X %*% fixef(m_glmm))
se_e <- sqrt(rowSums((X %*% as.matrix(vcov(m_glmm))) * X))

newd <- newd |>
  mutate(
    fit = plogis(eta),
    lo  = plogis(eta - 1.96 * se_e),
    hi  = plogis(eta + 1.96 * se_e),
    pO2c = pO2c_10 * PO2_UNIT
  )

ggplot(newd, aes(pO2c, fit, colour = Genotype, fill = Genotype)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.18, colour = NA) +
  geom_line(linewidth = 1.1) +
  geom_rug(data = filter(dat, isDying),
           aes(x = pO2c_calc, y = NULL), sides = "t",
           inherit.aes = FALSE, colour = COL_DYING, alpha = 0.9) +
  geom_rug(data = filter(dat, !isDying),
           aes(x = pO2c_calc, y = NULL), sides = "b",
           inherit.aes = FALSE, colour = COL_ALIVE, alpha = 0.25) +
  scale_colour_manual(values = c(WT = COL_WT, Tg = COL_TG)) +
  scale_fill_manual(values   = c(WT = COL_WT, Tg = COL_TG)) +
  labs(
    title = "Predicted probability that a neuron is dying",
    subtitle = "Population-average (u = 0) with 95% CI; rug = dying (top, red) and non-dying (bottom, grey)",
    x = expression(Delta*pO[2]~"from mouse mean (mmHg)"),
    y = "Pr(dying)"
  )
```

# 5. Rare-event and clustering safeguards

Three independent estimators. **Agreement across them is the evidence;
a result that appears in only one is not.**

## 5.1 Firth penalised logistic regression

Firth's penalty removes the small-sample bias of maximum likelihood and
gives finite estimates even under separation. Here `CaseID` enters as a
fixed effect — a *conditional* specification that removes each mouse's
baseline exactly rather than assuming a normal distribution over 7
mice.

Note that `Genotype` is nested within `CaseID`, so a genotype *main*
effect is not estimable alongside mouse fixed effects; it is absorbed.
The interaction *is* estimable because it is a within-mouse contrast.

```{r firth}
m_firth <- glm(
  isDying ~ CaseID + pO2c_10 + pO2c_10:Genotype,
  data   = dat,
  family = binomial("logit"),
  method = "brglmFit",       # brglm2
  type   = "AS_mean"         # = Firth's penalised likelihood
)

ci_firth <- suppressMessages(confint(m_firth))   # profile-likelihood CIs

firth_tbl <- tibble(
  term      = names(coef(m_firth)),
  OR        = exp(coef(m_firth)),
  conf.low  = exp(ci_firth[, 1]),
  conf.high = exp(ci_firth[, 2]),
  p.value   = coef(summary(m_firth))[, "Pr(>|z|)"]
) |>
  filter(grepl("pO2c_10", term))

kable(firth_tbl, digits = 3,
      caption = "Firth penalised estimates (mouse fixed effects absorbed; profile-likelihood CIs)")
```

`logistf::logistf(isDying ~ CaseID + pO2c_10 + pO2c_10:Genotype, data = dat, pl = TRUE)`
is an equivalent alternative if you already have that package installed;
`brglm2` is used here because it plugs directly into `glm()`.

## 5.2 GEE with mouse as the cluster

```{r gee}
dat_gee <- dat |> arrange(CaseID) |> mutate(CaseID_f = factor(CaseID))

m_gee <- geeglm(
  as.integer(isDying) ~ pO2c_10 * Genotype,
  id      = CaseID_f,
  data    = dat_gee,
  family  = binomial(link = "logit"),
  corstr  = "exchangeable"
)

gee_tbl <- broom::tidy(m_gee, conf.int = TRUE, exponentiate = TRUE) |>
  select(term, OR = estimate, conf.low, conf.high, p.value)

kable(gee_tbl, digits = 3, caption = "GEE, exchangeable working correlation, robust SE")
```

> **Caveat.** The sandwich variance estimator is anti-conservative with
> few clusters; the usual guidance is ~40+. With **7 mice** the GEE
> p-values are optimistic and should be read as a directional
> cross-check on the coefficient, not as inference.

## 5.3 Within-mouse permutation test (assumption-free)

The most trustworthy inference here. Under the null "dying status is
unrelated to pO~2~ *within* a mouse," we can permute the `isDying`
labels within each mouse and rebuild the null distribution directly.
This respects the clustering exactly, makes no distributional
assumption, and is unaffected by the small number of events.

Statistic: the mean within-mouse centred pO~2~ of dying neurons,
averaged over mice (mouse-weighted, so the 1,712-neuron mouse does not
swamp the others).

```{r permutation}
N_PERM <- 10000

perm_stat <- function(dying_vec, dat) {
  d <- dat
  d$dy <- dying_vec
  d |>
    group_by(Genotype, CaseID) |>
    filter(sum(dy) > 0) |>
    summarise(m = mean(pO2c_calc[dy]), .groups = "drop") |>
    group_by(Genotype) |>
    summarise(stat = mean(m), .groups = "drop")
}

obs <- perm_stat(dat$isDying, dat)

null_draws <- replicate(N_PERM, {
  shuffled <- dat |>
    group_by(CaseID) |>
    mutate(dy = sample(isDying)) |>
    ungroup() |>
    pull(dy)
  s <- perm_stat(shuffled, dat)
  c(WT = s$stat[s$Genotype == "WT"],
    Tg = s$stat[s$Genotype == "Tg"])
})

perm_res <- tibble(
  Genotype  = c("WT", "Tg"),
  observed  = c(obs$stat[obs$Genotype == "WT"], obs$stat[obs$Genotype == "Tg"]),
  null_mean = rowMeans(null_draws),
  null_sd   = apply(null_draws, 1, sd),
  p_two_sided = c(
    mean(abs(null_draws["WT", ]) >= abs(obs$stat[obs$Genotype == "WT"])),
    mean(abs(null_draws["Tg", ]) >= abs(obs$stat[obs$Genotype == "Tg"]))
  )
)

kable(perm_res, digits = 4,
      caption = paste0("Within-mouse permutation test, ", N_PERM,
                       " permutations. Observed = mean centred pO2 of dying neurons (mmHg), mouse-weighted."))
```

```{r fig-perm, fig.height=4}
tibble(
  Genotype = rep(c("WT", "Tg"), each = N_PERM),
  stat     = c(null_draws["WT", ], null_draws["Tg", ])
) |>
  mutate(Genotype = factor(Genotype, levels = c("WT", "Tg"))) |>
  ggplot(aes(stat, fill = Genotype)) +
  geom_histogram(bins = 60, colour = NA, alpha = 0.7) +
  geom_vline(data = perm_res |>
               mutate(Genotype = factor(Genotype, levels = c("WT", "Tg"))),
             aes(xintercept = observed),
             colour = COL_DYING, linewidth = 1) +
  scale_fill_manual(values = c(WT = COL_WT, Tg = COL_TG)) +
  facet_wrap(~ Genotype, scales = "free") +
  labs(
    title = "Permutation null vs observed",
    subtitle = "Red line = observed mean centred pO2 of dying neurons",
    x = expression("Mouse-weighted mean "*Delta*pO[2]*" of dying neurons (mmHg)"),
    y = "Permutations"
  )
```

## 5.4 Method comparison — forest plot

```{r fig-forest, fig.height=4.2}
comp <- bind_rows(
  or_glmm  |> filter(grepl("pO2c_10", term)) |> mutate(Method = "GLMM (glmer)"),
  firth_tbl                                   |> mutate(Method = "Firth (logistf)"),
  gee_tbl  |> filter(grepl("pO2c_10", term)) |> mutate(Method = "GEE (geepack)")
) |>
  mutate(
    term = recode(term,
                  "pO2c_10"                = "pO2 slope (WT)",
                  "pO2c_10:GenotypeTg"     = "pO2 slope difference (Tg - WT)"),
    Method = factor(Method, levels = c("GLMM (glmer)", "Firth (logistf)", "GEE (geepack)"))
  )

kable(comp |> select(Method, term, OR, conf.low, conf.high, p.value),
      digits = 3, caption = "All three estimators side by side")

ggplot(comp, aes(x = OR, y = Method, colour = term)) +
  geom_vline(xintercept = 1, linetype = 2, colour = "grey55") +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high),
                  position = position_dodge(width = 0.5), linewidth = 0.7) +
  scale_x_log10() +
  scale_colour_manual(values = c(COL_WT, COL_TG)) +
  facet_wrap(~ term, ncol = 1, scales = "free_y") +
  labs(
    title = "Odds ratio per 10 mmHg of within-mouse pO2 deviation",
    subtitle = "Consistency across estimators is the evidence",
    x = "Odds ratio (log scale)", y = NULL, colour = NULL
  ) +
  theme(legend.position = "none")
```

# 6. Sensitivity analyses

## 6.1 Within-mouse *z*-score instead of mmHg

Mice differ in pO~2~ *spread* as well as mean. Rescaling to within-mouse
SD units checks that the result is not driven by one wide-variance
animal.

```{r sens-z}
m_z <- glmer(isDying ~ z_calc * Genotype + (1 | CaseID),
             data = dat, family = binomial,
             control = glmerControl(optimizer = "bobyqa",
                                    optCtrl = list(maxfun = 2e5)))

broom.mixed::tidy(m_z, effects = "fixed", conf.int = TRUE,
                  exponentiate = TRUE) |>
  select(term, OR = estimate, conf.low, conf.high, p.value) |>
  kable(digits = 3, caption = "OR per 1 within-mouse SD of pO2")
```

## 6.2 Absolute pO~2~ (not recommended as primary)

Shown for completeness. With 7 mice the between-animal differences in
Figure 3.1 are only partially absorbed by the random intercept, so this
estimate blends within- and between-mouse variation.

```{r sens-raw}
m_raw <- glmer(isDying ~ pO2_10 * Genotype + (1 | CaseID),
               data = dat, family = binomial,
               control = glmerControl(optimizer = "bobyqa",
                                      optCtrl = list(maxfun = 2e5)))

broom.mixed::tidy(m_raw, effects = "fixed", conf.int = TRUE,
                  exponentiate = TRUE) |>
  select(term, OR = estimate, conf.low, conf.high, p.value) |>
  kable(digits = 3, caption = "Absolute pO2 model, OR per 10 mmHg")
```

## 6.3 Leave-one-mouse-out

With 7 animals, one mouse can drive the whole result. Refit dropping
each in turn.

```{r loo}
loo <- lapply(unique(dat$CaseID), function(cs) {
  d <- filter(dat, CaseID != cs)
  fit <- try(glmer(isDying ~ pO2c_10 * Genotype + (1 | CaseID),
                   data = d, family = binomial,
                   control = glmerControl(optimizer = "bobyqa",
                                          optCtrl = list(maxfun = 2e5))),
             silent = TRUE)
  if (inherits(fit, "try-error")) return(NULL)
  b <- fixef(fit); s <- sqrt(diag(as.matrix(vcov(fit))))
  tibble(
    dropped      = cs,
    OR_pO2_WT    = exp(b["pO2c_10"]),
    OR_interact  = exp(b["pO2c_10:GenotypeTg"]),
    p_interact   = 2 * pnorm(-abs(b["pO2c_10:GenotypeTg"] /
                                    s[names(b) == "pO2c_10:GenotypeTg"]))
  )
}) |> bind_rows()

kable(loo, digits = 3,
      caption = "Leave-one-mouse-out. Large swings mean the result rests on a single animal.")
```

# 7. Summary of findings

```{r summary-numbers}
summ <- tibble(
  Quantity = c(
    "WT slope, OR per 10 mmHg (GLMM)",
    "Tg slope, OR per 10 mmHg (GLMM)",
    "Interaction OR (Tg vs WT), GLMM",
    "Interaction OR, Firth",
    "Interaction OR, GEE",
    "LRT p, interaction",
    "LRT p, any pO2 effect",
    "Permutation p, WT",
    "Permutation p, Tg"
  ),
  Value = c(
    sprintf("%.2f [%.2f, %.2f]", slopes_tbl$OR[1], slopes_tbl$conf.low[1], slopes_tbl$conf.high[1]),
    sprintf("%.2f [%.2f, %.2f]", slopes_tbl$OR[2], slopes_tbl$conf.low[2], slopes_tbl$conf.high[2]),
    sprintf("%.2f", or_glmm$OR[or_glmm$term == "pO2c_10:GenotypeTg"]),
    sprintf("%.2f", firth_tbl$OR[grepl(":", firth_tbl$term)]),
    sprintf("%.2f", gee_tbl$OR[gee_tbl$term == "pO2c_10:GenotypeTg"]),
    sprintf("%.3f", anova(m_no_int, m_glmm)$`Pr(>Chisq)`[2]),
    sprintf("%.3f", anova(m_no_pO2, m_no_int)$`Pr(>Chisq)`[2]),
    sprintf("%.3f", perm_res$p_two_sided[perm_res$Genotype == "WT"]),
    sprintf("%.3f", perm_res$p_two_sided[perm_res$Genotype == "Tg"])
  )
)
kable(summ, caption = "Headline numbers, auto-filled from the fitted models")
```

**Reading these data.** Fill in the narrative once you have run the
document, but the structure of the answer is:

1. **Genotype effect on death rate** — the `GenotypeTg` term. This is
   the one term with a large, precisely estimated effect, and it is
   what you would expect: rTg4510 neurons die at a much higher rate.
   Treat its confidence interval with caution — it rests on 4 vs 3
   animals, so the *precision* is overstated even though the direction
   is clear.
2. **pO~2~ effect within mice** — the `pO2c_10` slope. An OR near 1
   with a CI spanning it means these data do not show dying neurons
   sitting in relatively hypoxic tissue *within* an animal.
3. **The interaction** — whether the pO~2~–death relationship differs
   by genotype. Compare the GLMM, Firth, GEE, and permutation results
   in §5.4; only a finding that survives all four is worth reporting.
4. **Power** — with `r sum(dat$isDying)` events, the CIs in §4.3 tell
   you what effect sizes these data could have detected. If the
   interval spans, say, 0.5 to 1.9, the study is compatible with a
   two-fold effect in either direction, and "no significant difference"
   should be written as "underpowered to detect", not "no effect".

# 8. Reproducibility

```{r session}
sessionInfo()
```
